Assessing the Replicability of Psychological Science Through Pedagogy

### 2016-08-22 paper/paper.Rmd
pipeline figure and other fig tweaking

```{r global_options, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=FALSE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(xtable)
library(readr)
library(tidyr)
library(ggplot2)
library(BayesFactor)
library(ggrepel)
#devtools::install_github("langcog/langcog")
library(langcog)
library(cowplot)
```

@@

All materials and methods for our replication studies are available at [https://github.com/StanfordPsych254](). All experiments were coded in JavaScript, HTML, and CSS so as to be run in a standard web browser. All analyses were written using R, a free and open-source platform for statistical analysis.

@@

We next turn to an assessment of the replication $p$-value to determine whether the key statistic found a significant effect at the traditional $p=.05$ level. Only `r sum(d$rep_p_value_05)` of the studies yielded a significant replication $p$-value. The well-documented "dance of the p-values" [@cumming2013] suggests that the inference drawn from a single replication p-value may not be  informative, however, and other metrics such as effect sizes and Bayes factors may provide a more nuanced perspective. While we preregistered an analysis determining whether the original reported effect size is included in the 95\% confidence interval found in the replication, we instead opted for a straightforward comparison of point estimates. This choice was motivated by recent discussions of the utility and validity of interpretating confidence intervals in this manner [@morey2016], as well as discrepencies in reporting effect sizes across the articles we included.

### Bayes factor and effect size analysis

Next, following @etz2016, we compared Bayes Factors between the original and replication. The Bayes Factor is an alternative to classical null hypothesis significance testing; it quantifies the evidence in favor of a hypothesis as the ratio of the likelihood of the data under that hypothesis relative to another [@jeffreys1961]. Its proponents point to many advantages over classical $p$-values. For example, it does not privilege the null hypothesis and can provide evidence either *for* or *against* it [see also @scheibehenne2016, @wagenmakers2016]. While the Bayes Factor often agrees with the $p$-value as to which hypothesis is more likely [@wetzels2011], they often substantially disagree on the strength of the evidence.

```{r}
# Note that the sofer paper is off the charts (d > 16) and excluded
p1 <- ggplot(d,aes(x=original_ES_d,y=rep_es)) +
  geom_point(alpha = .8, 
             aes(color=replicated_TA_factor, size = collected_n)) +
  geom_text_repel(aes(label = first_author_name), size = 2, force = 2) +
  scale_x_continuous(limits = c(0, 3.5), 
                breaks = c(.2, .5, .8, 1, 2, 3)) +
  scale_y_continuous(limits = c(0, 3.5), 
                breaks = c(.2, .5, .8, 1, 2, 3)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_smooth(method = "lm", se=FALSE) +
  xlab("Orig. effect size (d)") + 
  ylab("Rep. effect size (d)") + 
  theme_bw(base_size=10) + theme(legend.position="bottom") + 
  scale_size(name = "Rep. N") +
  scale_color_solarized(name = "Subjective rep. rating")

@@

p2 <- ggplot(tstatDF, aes(x = origBF, y = repBF)) +
  geom_point(alpha = .8, aes(size = df_ratio, 
                             col = replicated_TA_factor)) +
  geom_abline(slope = 1, lty = 2) + 
  geom_smooth(method = "lm", se=FALSE) +
  geom_text_repel(aes(label = first_author_name), size = 2, force = 2) +
  xlab("Orig. Bayes factor") + 
  ylab("Rep. Bayes factor") + 
  scale_x_log10(limits = c(10**-1.75, 10**11), 
                breaks = c(1/10, .3, 3, 10, 100, 1e+3, 1e+6, 1e+9),
                labels = c("1e-1", "3e-1", "3", "1e1","1e2", "1e3", 
                           "1e6","1e9")) +
  scale_y_log10(limits = c(10**-1.75, 10**11),
                breaks = c(1/10, .3, 3, 10, 100, 1e+3, 1e+6, 1e+9),
                labels = c("1e-1", "3e-1", "3", "1e1","1e2", "1e3", 
                           "1e6","1e9")) + 
  theme_bw(base_size=10) + theme(legend.position="bottom") + 
  scale_size_continuous(name = "Ratio of rep. df to orig. df") + 
  scale_color_solarized(name = "Subjective rep. rating")

@@

For each project, we present a summary of the finding, the key statistical test, and the outcomes of the replication. Side-by-side visualizations are presented in Figure SM1.

>>>>

```{r global_options, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=FALSE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(xtable)
library(readr)
library(tidyr)
library(ggplot2)
library(BayesFactor)
library(ggrepel)
#devtools::install_github("langcog/langcog")
library(langcog)
library(cowplot)
library(png)
```

@@

All materials and methods for our replication studies are available at [https://github.com/StanfordPsych254](). All experiments were coded in JavaScript, HTML, and CSS so as to be run in a standard web browser. All analyses were written using R, a free and open-source platform for statistical analysis. A schematic of our class workflow is shown in Figure 1.

```{r, fig.cap="A schematic view of our class timeline for replication projects. R indicates the approximate timing of instructor team reviews of student materials.", fig.height=4, fig.width=6, fig.align="center", fig.pos="t"}

img <- readPNG(source = "../figures/pipeline.png")
plot.new()
plot.window(0:1, 0:1)

#fill plot with image
usr<-par("usr")    
rasterImage(img, usr[1], usr[3], usr[2], usr[4])
```

@@

We next turn to an assessment of the replication $p$-value to determine whether the key statistic found a significant effect at the traditional $p=.05$ level. Only `r sum(d$rep_p_value_05)` of the studies yielded a significant replication $p$-value. The well-documented "dance of the p-values" [@cumming2013] suggests that the inference drawn from a single replication *p*-value may not be  informative, however, and other metrics such as effect sizes and Bayes factors may provide a more nuanced perspective. While we preregistered an analysis determining whether the original reported effect size is included in the 95\% confidence interval found in the replication, we instead opted for a straightforward comparison of point estimates. This choice was motivated by recent discussions of the utility and validity of interpretating confidence intervals in this manner [@morey2016], as well as discrepencies in reporting effect sizes across the articles we included (which made computing confidence intervals difficult in some cases).

### Effect size and Bayes Factor analysis

We first compared effect sizes from the original and replication studies (Figure 2A). Effect sizes were highly correlated between the two sets of studies 
(r = `r signif(cor.test(d$original_ES_d,d$rep_es)$estimate, digits =2)`, p < .0001), but they were generally smaller in the replications 
(`r signif(coef(lm(d$rep_es~d$original_ES_d-1))[1], digits = 2)`\% of the original [`r signif(confint(lm(d$rep_es~d$original_ES_d-1))[1], digits = 2)` - `r signif(confint(lm(d$rep_es~d$original_ES_d-1))[2], digits = 2)`]). This estimate should be viewed with caution, however, due to its dependence on one very large effect size estimate ($d > 15$) from a single study.

Next, following @etz2016, we compared Bayes Factors between the original and replication (Figure 2B). The Bayes Factor is an alternative to classical null hypothesis significance testing; it quantifies the evidence in favor of a hypothesis as the ratio of the likelihood of the data under that hypothesis relative to another [@jeffreys1961]. Its proponents point to many advantages over classical $p$-values. For example, it does not privilege the null hypothesis and can provide evidence either *for* or *against* it [see also @scheibehenne2016, @wagenmakers2016]. While the Bayes Factor often agrees with the $p$-value as to which hypothesis is more likely [@wetzels2011], they often substantially disagree on the strength of the evidence.

```{r, fig.cap="Replication effect size (A) and Bayes factor (B), plotted by the original effect size and Bayes factor, respectively. Point size shows replication N (A) and ratio of test degrees of freedom (B), color indicates subjective replication assessments by the authors. Note that the Sofer et al. (2015) replication showed ES and BF too large to plot easily on these axes and is hence excluded.", fig.pos="ht", fig.height=4}

# Note that the sofer paper is off the charts (d > 16) and excluded
p1 <- ggplot(d,aes(x=original_ES_d,y=rep_es)) +
  geom_point(alpha = .8, 
             aes(color=replicated_TA_factor, size = collected_n)) +
  geom_text_repel(aes(label = first_author_name), size = 2, force = 2) +
  scale_x_continuous(limits = c(0, 3.5), 
                breaks = c(.2, .5, .8, 1, 2, 3)) +
  scale_y_continuous(limits = c(0, 3.5), 
                breaks = c(.2, .5, .8, 1, 2, 3)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_smooth(method = "lm", se=FALSE) +
  xlab("Orig. effect size (d)") + 
  ylab("Rep. effect size (d)") + 
  theme_bw(base_size=10) + theme(legend.position="bottom") + 
  scale_size(name = "Rep. N") +
  # scale_color_solarized(guide = FALSE)
  scale_color_solarized(name = "Subjective rep. rating")

@@

p2 <- ggplot(tstatDF, aes(x = origBF, y = repBF)) +
  geom_point(alpha = .8, aes(size = df_ratio, 
                             col = replicated_TA_factor)) +
  geom_abline(slope = 1, lty = 2) + 
  geom_smooth(method = "lm", se=FALSE) +
  geom_text_repel(aes(label = first_author_name), size = 2, force = 2) +
  xlab("Orig. Bayes Factor") + 
  ylab("Rep. Bayes Factor") + 
  scale_x_log10(limits = c(10**-1.75, 10**11), 
                breaks = c(1/10, .3, 3, 10, 100, 1e+3, 1e+6, 1e+9),
                labels = c("1e-1", "3e-1", "3", "1e1","1e2", "1e3", 
                           "1e6","1e9")) +
  scale_y_log10(limits = c(10**-1.75, 10**11),
                breaks = c(1/10, .3, 3, 10, 100, 1e+3, 1e+6, 1e+9),
                labels = c("1e-1", "3e-1", "3", "1e1","1e2", "1e3", 
                           "1e6","1e9")) + 
  theme_bw(base_size=10) + theme(legend.position="bottom") + 
  scale_size_continuous(name = "Ratio of rep. df to orig. df") + 
  scale_color_solarized(name = "Subjective rep. rating")

@@

For each project, we present a summary of the finding, the key statistical test, and the outcomes of the replication.
